---
title: "Bethnic Communities Initial Data Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(corrplot)
library(ggplot2)
```

Insert Our Data Set
```{r}
getwd()
#?setwd("C:/Users/This PC/Documents")
bethniccom <- read.csv("Shannon Wiener Merged.csv")
```
```{r}
View(bethniccom)
```

First Analysis Compare How Growth of Organisms Impacts the Growth of Other Organisms

```{r}
bethniccom_onlyorg <- bethniccom[,-c(1,2,3,21)]

bethniccom_onlyorg <- bethniccom_onlyorg/100
#View(bethniccom_onlyorg)
```
arbarescent bryozoans
```{r}
cp <- cor(bethniccom_onlyorg)

corrplot(cp,method = "circle")

#bethniccom <- bethniccom %>%
  #rename(Dock = ï..Dock)
```
```{r}
dockcol <- as.factor(bethniccom$Dock)
matcol <- as.factor(bethniccom$Mat)
shellcol <- as.factor(bethniccom$Shell)
seasoncol <- as.factor(bethniccom$Season)

bethniccom_new <- cbind(dockcol, matcol, shellcol, bethniccom_onlyorg, seasoncol)
cp1 <- cor(bethniccom_new)

corrplot(cp1)
```

```{r}
cp
```

```{r}
res2 <- rcorr(as.matrix(cp))
res2

```

```{r}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```
```{r}
res2df <- flattenCorrMatrix(res2$r, res2$P)
res2df
#View(res2df)
max(abs(res2df$p))
```

The highest p value is between Ribbed Mussel and Tubeworms. The lowest p value is between Barnacles and Oysters.



Scatterplot of Barnacles vs. All Other Organisms
```{r}
co <- c(colnames(bethniccom_onlyorg))
View(co)

```
```{r}
ggplot(data = bethniccom_onlyorg, aes(Barnacles,Arb..Bryo.)) + geom_point() + geom_smooth() + theme_bw()
```



, color = "blue"
, color = "blue"
, color = "blue"
, color = "red"
, color = "dark green"
, color = "dark green"
, color = "darkpurple"
, color = "darkpurple"
, color = "darkorange"
, color = "darkorange"
, color = "yellow"
, color = "yellow"
, color = "darkcyan"
, color = "darkcyan"
, color = "darkolivegreen1"
, color = "darkolivegreen1"
, color = "deeppink"
, color = "deeppink"
, color = "gray0"
, color = "gray0"
, color = "khaki1"
, color = "khaki1"
, color = "lightsalmon"
, color = "lightsalmon"
```{r}

colors <- c("Arb..Bryo." = "blue", "Enc..Bryo." = "red", "Tubeworms" = "dark green","Hydroids" = "purple", "Tunicates" = "darkorange","Oysters" = "yellow","Sponges" = "darkcyan", "Anemones" = "darkolivegreen1", "Algae" = "deeppink","Biofilm" = "gray0","Slippersnail" = "khaki1","Feather.Duster.Worm" = "lightsalmon")

#the lines of best fit on this graph are generated using the Loess Method
ggplot(data = bethniccom_onlyorg, mapping = aes(x = Barnacles)) +
  geom_point(aes(y = Arb..Bryo., color = "Arb..Bryo.")) +
  geom_smooth(aes(y = Arb..Bryo., color = "Arb..Bryo.")) +
  geom_point(aes(y = Enc..Bryo., color = "Enc..Bryo.")) +
  geom_smooth(aes(y = Enc..Bryo., color = "Enc..Bryo.")) +
  geom_point(aes(y = Tubeworms, color = "Tubeworms")) +
  geom_smooth(aes(y = Tubeworms, color = "Tubeworms")) +
  geom_point(aes(y = Hydroids, color = "Hydroids")) +
  geom_smooth(aes(y = Hydroids, color = "Hydroids")) +
   geom_point(aes(y = Tunicates, color = "Tunicates")) +
  geom_smooth(aes(y = Tunicates, color = "Tunicates")) +
  geom_point(aes(y = Oysters, color = "Oysters")) +
  geom_smooth(aes(y = Oysters, color = "Oysters")) +
  geom_point(aes(y = Sponges, color = "Sponges")) +
  geom_smooth(aes(y = Sponges, color = "Sponges")) +
  geom_point(aes(y = Anemones, color = "Anemones")) +
  geom_smooth(aes(y = Anemones, color = "Anemones")) +
  geom_point(aes(y = Algae, color = "Algae")) +
  geom_smooth(aes(y = Algae, color = "Algae")) +
  geom_point(aes(y = Biofilm, color = "Biofilm")) +
  geom_smooth(aes(y = Biofilm, color = "Biofilm")) +
  geom_point(aes(y = Slippersnail, color = "Slippersnail")) +
  geom_smooth(aes(y = Slippersnail, color = "Slippersnail")) +
  geom_point(aes(y = Feather.Duster.Worm, color = "Feather.Duster.Worm")) +
  geom_smooth(aes(y = Feather.Duster.Worm, color = "Feather.Duster.Worm")) +
  #geom_point(aes(y = Amphipod.Tubes, shape = 4)) +
  #geom_smooth(aes(y = Amphipod.Tubes, color = "ivory")) +
  ylab("Organism Type") + 
  labs(color = "Legend") +
  ggtitle("Percent Cover of Other Organisms with respect to Barnacle Cover") +
  scale_color_manual(values = colors)
```

























