---
title: "Fall vs. Spring Bethnic Com"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(data.table)
library(dplyr)
library(Rmisc)
```


```{r}

getwd()
#setwd("C:/Users/This PC/Documents")
bethniccom <- read.csv("Shannon Wiener Merged Fixed All Data.csv",header = TRUE)

```
```{r}
View(bethniccom)
#rm(bethniccom)
```



```{r}
bethniccom_onlyorg <- bethniccom[,-c(1,2,3,21)]

bethniccom_onlyorg <- bethniccom_onlyorg/100
#View(bethniccom_onlyorg)
```

```{r}
#bethniccom <- bethniccom %>%
  #rename(dock = Dock)

View(bethniccom)
```


```{r}
names(bethniccom)
```




First we will sort the data so we can compare organisms for fall vs. spring for individual shells
```{r}
#sortbc <- bethniccom %>%
 # mutate(id = gsub(" ", "",paste(Dock, Mat, Shell),fixed = TRUE)) %>%
  #arrange(id) 
  #arrange(Shell,Mat)
#View(sortbc)



```

Now do this for individual docks
```{r}
sortbc_A1A <- sortbc %>%
  filter(Dock == "A1A")
#View(sortbc_A1A)
```

```{r}
sortbc_Aquarina <- sortbc %>%
  filter(Dock == "Aquarina")
#View(sortbc_Aquarina)
```
```{r}
sortbc_LighthouseCove <- sortbc %>%
  filter(Dock == "Lighthouse Cove")
#View(sortbc_LighthouseCove)
```
```{r}
sortbc_IAP <- sortbc %>%
  filter(Dock == "IAP")
#View(sortbc_IAP)
```
```{r}
sortbc_MBP <- sortbc %>%
  filter(Dock == "MBP")
#View(sortbc_MBP)
```
```{r}
sortbc_MelbourneShores <- sortbc %>%
  filter(Dock == "Melbourne Shores")
#View(sortbc_MelbourneShores)
```
```{r}
sortbc_Sebastian <- sortbc %>%
  filter(Dock == "Sebastian")
#View(sortbc_Sebastian)
```
```{r}
sortbc_Wingate <- sortbc %>%
  filter(Dock == "Wingate")
#View(sortbc_Wingate)
```


Now we will create a histogram of organism levels in fall vs. spring per dock

Start with A1A
```{r, error=TRUE}
#View(sortbc_A1A)

A1Apair <- sortbc_A1A %>%
  filter(id == sortbc_A1A$id[1])


#View(A1Apair)
A1Apairdf <- data.frame(t(A1Apair)) 
A1Apairdf <- A1Apairdf[-c(1,2,3,21,22),]


#View(A1Apairdf)

setDT(A1Apairdf, keep.rownames = TRUE)[]


names(A1Apairdf)<-c("Organism","Fall","Spring")

A1Apairdf


```

Trying to make a barplot

```{r, error=TRUE}


for (i in 1:17) {
  A1Apairdf$Fall[i] <- str_squish(A1Apairdf$Fall[i])
}

View(A1Apairdf)

A1Apairdf[,2] <- as.factor(A1Apairdf[,2])
A1Apairdf[,3] <- as.factor(A1Apairdf[,3])


ggplot(data = A1Apairdf, aes(x = factor(Organism),y=Fall)) + geom_col(position = "dodge2") + theme_bw()
ggplot(data = A1Apairdf, aes(x = Organism, y = Spring)) + geom_bar(stat = "identity") 

?geom_bar()
```

More Barplot Attempts

```{r, error=TRUE}
barplot(as.numeric(A1Apairdf$Organism),A1Apairdf$Fall)
```

```{r}
?mutate()
colnames(bethniccom)
?pivot_longer()
```

Time to try something else, going back to the original data set and trying to organize it

```{r}
bethniccom_practice <- bethniccom[,4:21]
bethniccom_practice <-bethniccom_practice %>% 
  pivot_longer(cols = !Season, names_to = "Organism",values_to = "PercentCover")
  
View(bethniccom_practice)
```

It worked! :) Now that we have organized data, we can start analysis after we get a starting understanding of the data.

First we have to change the column types.

```{r}
mode(bethniccom_practice)
bethniccom_practice$Season <- as.factor(bethniccom_practice$Season)
bethniccom_practice$Organism <- as.factor(bethniccom_practice$Organism)
bethniccom_practice$PercentCover <- as.numeric(bethniccom_practice$PercentCover)
str(bethniccom_practice)
```

To get an understanding of how the data is distributed, let's make a boxplot of organism percent conver by season.

```{r}
ggplot(bethniccom_practice, aes(x = Organism, y = PercentCover, color = Season)) + geom_boxplot() + ggtitle("Fall vs. Spring Organism Distribution") + labs(x = "Organism",y = "Percent Cover") + coord_flip() + theme_bw()
```
Now we will try and do some analysis by getting a summary of the data and making boxplots of the confidence intervals.
```{r}
sumbc <- summarySE(bethniccom_practice,
                measurevar="PercentCover",
                groupvars=c("Season","Organism"))

#sumbc
pd = position_dodge(.2)

ggplot(sumbc, aes(x=Organism,
                y=PercentCover,
                color=Season)) +
  geom_errorbar(aes(ymin=PercentCover-se,
                    ymax=PercentCover+se),
                width=.2, size=0.7, position=pd) +
  geom_point(shape=15, size=4, position=pd) +
  theme(
    axis.title.y = element_text(vjust= 1.8),
    axis.title.x = element_text(vjust= -0.5),
    axis.title = element_text(face = "bold")) +
  scale_color_manual(values = c("red", "blue")) + 
  ggtitle("Confidence Intervals") +
  coord_flip() + theme_bw() + labs(x = "Organism",y = "Percent Cover")
```

Now we will do ANOVA for Percent Cover Based on Organism
```{r}
attach(bethniccom_practice)
levels(Organism)
plot(PercentCover~Organism)
anvml_pco <- aov(PercentCover~Organism)
summary(anvml_pco)

pairwise.t.test(PercentCover, Organism, p.adjust="bonferroni")
TukeyHSD(anvml_pco, conf.level = 0.95)
```

Now we will do Anova for Percent Cover based on Season
```{r}
levels(Season)
plot(PercentCover~Season)
anvml_pcs <- aov(PercentCover~Season)
summary(anvml_pcs)

pairwise.t.test(PercentCover, Season, p.adjust="bonferroni")
TukeyHSD(anvml_pcs, conf.level = 0.95)
```
Anova of Season and Organism for Percent Cover
```{r}
plot(PercentCover~Season+Organism)

anvml_pcso <- aov(PercentCover~Season+Organism)
summary(anvml_pcso)


pairwise.t.test(PercentCover, Season, p.adjust="bonferroni")
TukeyHSD(anvml_pcso, conf.level = 0.95)

#View(bethniccom)
```

Let's compare which model out of our 3 anovas is better

```{r}
library(AICcmodavg)

model.set <- list(anvml_pco,anvml_pcs,anvml_pcso,anvml_pcosi,anvml_pcoi)
model.names <- c("Anova Organism", "Anova Season", "Anova Both","Anova Both and Interaction","Anova Organism and Interaction")

aictab(model.set, modnames = model.names)
?aictab()
```


```{r}
#par(mfrow=c(2,2))
par(mfrow=c(1,1))
plot(anvml_pco)

```

Now we will do Anova for Percent Cover based on Season
```{r}
levels(Season)
plot(PercentCover~Season+Organism+Season*Organism)
anvml_pcosi <- aov(PercentCover~Season+Organism+Season*Organism)
summary(anvml_pco)

pairwise.t.test(PercentCover, Season, p.adjust="bonferroni")
TukeyHSD(anvml_pcs, conf.level = 0.95)
```
```{r}
#par(mfrow=c(2,2))
par(mfrow=c(1,1))
plot(anvml_pcoi)

```

```{r}
attach(bethniccom_practice)
levels(Season)
plot(PercentCover~Organism+Season*Organism)
anvml_pcoi <- aov(PercentCover~Organism+Season*Organism)
summary(anvml_pcoi)


#pairwise.t.test(PercentCover, Season, p.adjust="bonferroni")
#TukeyHSD(anvml_pcs, conf.level = 0.95)
```

```{r}
#par(mfrow=c(2,2))
par(mfrow=c(1,1))
plot(anvml_pcoi)

```

Now we will do log transformations
```{r}
library(openintro)
library(dplyr)
library(ggplot2)
library(e1071)
```
Assess Skewedness of data
```{r}
skewness(bethniccom_practice$PercentCover)
```
This indicates that the data is highly skewed positive. A visual representation is availible in the following histogram

This is the data using log and log 10 without adding 1
```{r}
ggplot(data = bethniccom_practice, aes(PercentCover)) + geom_histogram()


bethniccom_logpractice <- cbind(bethniccom_practice, log(bethniccom_practice$PercentCover),log10(bethniccom_practice$PercentCover))

#View(bethniccom_logpractice)
str(bethniccom_logpractice)

names(bethniccom_logpractice)[names(bethniccom_logpractice) == "log(bethniccom_practice$PercentCover)"] <- "logPercentCover"
names(bethniccom_logpractice)[names(bethniccom_logpractice) == "log10(bethniccom_practice$PercentCover)"] <- "log10PercentCover"

ggplot(data = bethniccom_logpractice, aes(x = logPercentCover)) +
  geom_histogram() + ggtitle("Percent Cover Log Transformation")
ggplot(data = bethniccom_logpractice, aes(x = log10PercentCover)) +
  geom_histogram() + ggtitle("Percent Cover Log10 Transformation")
```

This is the data using 1 + log and log 10
```{r}
ggplot(data = bethniccom_practice, aes(PercentCover)) + geom_histogram()+ ggtitle("Percent Cover")


bethniccom_logpractice <- cbind(bethniccom_practice, log(1 + bethniccom_practice$PercentCover),log10(1 + bethniccom_practice$PercentCover))

#View(bethniccom_logpractice)
str(bethniccom_logpractice)

names(bethniccom_logpractice)[names(bethniccom_logpractice) == "log(1 + bethniccom_practice$PercentCover)"] <- "logPercentCover"
names(bethniccom_logpractice)[names(bethniccom_logpractice) == "log10(1 + bethniccom_practice$PercentCover)"] <- "log10PercentCover"


ggplot(data = bethniccom_logpractice, aes(x = PercentCover)) +
  geom_histogram() + ggtitle("Percent Cover No Transformation")
ggplot(data = bethniccom_logpractice, aes(x = logPercentCover)) +
  geom_histogram() + ggtitle("Percent Cover Log Transformation")
ggplot(data = bethniccom_logpractice, aes(x = log10PercentCover)) +
  geom_histogram() + ggtitle("Percent Cover Log10 Transformation")
```

